### 2025-04-18 methodological_analysis.qmd
Add files via upload

---
title: "ZIPLN-network analysis"
format: html
editor: visual
---

# General questions

We aim at using the ZIPLN-network framework to answer the following questions:

1.  Does accounting for the zero-inflation in the data improve the results of the model ? How so ?

2.  Are the two datasets' distributions significantly different ?

3.  How do the environmental covariates influence the results ? Which ones should we include ?

4.  How robust are the different association to changes in the model (inclusion of different covariates, zero-inflation modelling...) ?

5.  Does including sampling months / years change the results for the Northern Range dataset ?

# Useful libraries & functions

```{r loading_libraries}
library(dplyr)
library(PLNmodels)
library(igraph)
library(FactoMineR)
library(ggplot2)
library(Metrics)
library(Hmsc)
library(corrplot)
set.seed(1)
```

```{r robustness}

robustness = function(models, models_names = NULL,
                      matrices_input = FALSE){
  N <- length(models) ; res <- matrix(rep(0, N * N), nrow = N) ; diag(res) <- 1
  for(i in 1:(N-1)){
    if(!matrices_input){
      mod1 <- models[[i]] ; net1   <- as.matrix(mod1$latent_network("partial_cor"))
      diag(net1) <- 0
    }else{net1 <- models[[i]]}
    count1 <- sum(net1 != 0)
    for(j in ((i + 1):N)){
      if(!matrices_input){
        mod2 <- models[[j]] ; net2   <- as.matrix(mod2$latent_network("partial_cor"))
        diag(net2) <- 0
      }else{net2 <- models[[j]]}
      count2 <- sum(net2 != 0)
      intersection <- sum(sign(net1) * sign(net2) == 1)
      res[i, j]    <- round(intersection / count1, 2)
      res[j, i]    <- round(intersection / count2, 2)
    }
  }
  if(!is.null(models_names)){
    rownames(res) <- models_names ; colnames(res) <- models_names
  }
  cat("M[i,j] gives the proportion of interactions contained in model i that are also in model j. \n")
  return(res)
}
```

```{r plot-network}
plot_network = function(net             ,
                        type = "partial_cor",            
                        edge.color      = c("#F8766D", "#00BFC4"),
                        remove.isolated = FALSE,
                        node.labels     = NULL,
                        layout          = layout_in_circle) {
  G <-  graph_from_adjacency_matrix(net, mode = "undirected", weighted = TRUE, diag = FALSE)

  if (!is.null(node.labels)) {
    igraph::V(G)$label <- node.labels
  } else {
    igraph::V(G)$label <- colnames(net)
  }
  ## Nice nodes
  V.deg <- degree(G)/sum(degree(G))
  igraph::V(G)$label.cex <- V.deg / max(V.deg) + .5
  igraph::V(G)$size <- V.deg * 100
  igraph::V(G)$label.color <- rgb(0, 0, .2, .8)
  igraph::V(G)$frame.color <- NA
  ## Nice edges
  igraph::E(G)$color <- ifelse(igraph::E(G)$weight > 0, edge.color[1], edge.color[2])
  if (type == "support")
    igraph::E(G)$width <- abs(igraph::E(G)$weight)
  else
    igraph::E(G)$width <- 15*abs(igraph::E(G)$weight)

  if (remove.isolated) {
    G <- delete.vertices(G, which(degree(G) == 0))
  }
  plot(G, layout = layout, font.size = 8)
  invisible(G)
}

```

# Data importation

### Regional dataset

```{r, loading regional dataset}
load("REGIONAL_DATASET_COUNTS.Rdata")
load("REGIONAL_DATASET_COVARIATES.Rdata")
```

```{r, formatting regional data for PLNmodels}
regional_data <- prepare_data(counts     = REGIONAL_DATASET_COUNTS,
                              covariates = REGIONAL_DATASET_COVARIATES)
```

### Northern Range dataset

```{r, loading Northern Range dataset}
load("NORTHERN_RANGE_DATASET_COUNTS.Rdata")
load("NORTHERN_RANGE_DATASET_COVARIATES.Rdata")
```

```{r, formatting Northern Range data for PLNmodels}
northern_range_data <- prepare_data(counts     = NORTHERN_RANGE_DATASET_COUNTS,
                                    covariates = NORTHERN_RANGE_DATASET_COVARIATES)
```

# Data visualization

We use log-scale heatmaps to visualize the data. \## Regional dataset

```{r regional_heatmap}
pheatmap::pheatmap(log(1 + regional_data$Abundance),
                   treeheight_row = 0, treeheight_col = 0, show_rownames = FALSE)
```

## Northern Range dataset

```{r NR_heatmap}
pheatmap::pheatmap(log(1 + northern_range_data$Abundance),
                   treeheight_row = 0, treeheight_col = 0, show_rownames = FALSE)
```

# Preparing the covariates

Both datasets have the following environmental covariates in common: width, depth, temperature, oxygen level and pH. Given the results, we choose to keep the first three principal components.

## PCA on the covariates

### Regional dataset

```{r standardizing_covariates_regional}
regional_covar <- regional_data %>% select(c("WIDTHavg", "DEPTHavg",
                                             "TEMPavg", "OXYGENavg",
                                             "pHaverage"))
regional_covar <- as.data.frame(scale(regional_covar))
```

```{r pca_regional}
pca_regional <- PCA(regional_covar)
print(pca_regional$eig)
```

Given the PCA results, the site indexed 40 seems to be an outlier. We remove it before further analysis.

```{r}
REGIONAL_DATASET_COUNTS     <- REGIONAL_DATASET_COUNTS[-c(40),]
REGIONAL_DATASET_COVARIATES <- REGIONAL_DATASET_COVARIATES[-c(40),]
regional_data <- prepare_data(counts     = REGIONAL_DATASET_COUNTS,
                              covariates = REGIONAL_DATASET_COVARIATES)
```

```{r}
pca_covar_regional <- data.frame("PC1" = unname(pca_regional$ind$coord[, 1]),
                                 "PC2" = unname(pca_regional$ind$coord[, 2]),
                                 "PC3" = unname(pca_regional$ind$coord[, 3]))
pca_covar_regional <- pca_covar_regional[-c(40),]
rownames(pca_covar_regional) <- NULL
```

```{r}
regional_data_pc <- prepare_data(counts = REGIONAL_DATASET_COUNTS,
                                 covariates = pca_covar_regional)
```

### Northern Range dataset

```{r standardizing_covariates_nr}
northern_range_covar <- northern_range_data %>%
                        select(c("width", "depth", "temperature", "O2", "pH"))
northern_range_covar <- as.data.frame(scale(northern_range_covar))
```

```{r pca_northern_range}
pca_northern_range <- PCA(northern_range_covar)
print(pca_northern_range$eig)
```

```{r}
pca_covar_northern_range <- data.frame("PC1" = unname(pca_northern_range$ind$coord[, 1]),
                                       "PC2" = unname(pca_northern_range$ind$coord[, 2]),
                                       "PC3" = unname(pca_northern_range$ind$coord[, 3]))
rownames(pca_covar_northern_range) <- rownames(NORTHERN_RANGE_DATASET_COUNTS)
```

```{r}
northern_range_data_pc <- prepare_data(counts     = NORTHERN_RANGE_DATASET_COUNTS,
                                       covariates = pca_covar_northern_range)
```

## Trying to get the altitude
```{r}
library(elevatr)
library(sf)
library(terra)
```

```{r}
# point <- st_as_sf(data.frame(x = -61.31968, y = 10.70517),
#                   coords = c("x", "y"), crs = 4326)
# 
# # Get elevation
# elevation <- get_elev_point(point, src = "aws")
# print(elevation)

altitude <- unlist(lapply(1:304, F = function(i){
  point <- st_as_sf(data.frame(x = northern_range_data$LONGITUDE[[i]], y = northern_range_data$LATITUDE[[i]]),
                  coords = c("x", "y"), crs = 4326)
  return(get_elev_point(point, src = "aws")$elevation)
}))

```

```{r}
get_elev_point(locations = mts, prj = ll_prj)
```

# PLN and ZIPLN (no network penalty) to compare the fitting qualities

## PLN - PC1 + PC2

### regional dataset

```{r}
regional_no_network <- PLN(Abundance ~ 1 + PC1 + PC2,
                             data = regional_data_pc)
```

```{r}
Y_real   <- regional_data_pc$Abundance
Y_fitted <- regional_no_network$fitted ; colnames(Y_fitted) <- colnames(Y_real)
regional_rmse <- rmse(Y_real, Y_fitted)
plot(Y_real, Y_fitted) ; print(regional_rmse)  
```

### NR dataset

```{r}
northern_range_no_network <- PLN(Abundance ~ 1 + PC1 + PC2,
                                 data = northern_range_data_pc)
```

```{r}
Y_real   <- northern_range_data_pc$Abundance
Y_fitted <- northern_range_no_network$fitted ; colnames(Y_fitted) <- colnames(Y_real)
northern_range_rmse <- rmse(Y_real, Y_fitted)
plot(Y_real, Y_fitted) ; print(northern_range_rmse)  
```

```{r}
fitting_results <- data.frame("Model" = c("PLN_PC1_PC2"),
                              "regional_RMSE" = c(round(regional_rmse, 3)),
                              "northern_range_RMSE" = c(round(northern_range_rmse, 3)))
```

## ZIPLN - PC1 + PC2

### regional dataset

```{r}
regional_no_network_ZI <- ZIPLN(Abundance ~ 1 + PC1 + PC2 | 1,
                                data = regional_data_pc)
```

```{r}
Y_real   <- regional_data_pc$Abundance
Y_fitted <- regional_no_network_ZI$fitted ; colnames(Y_fitted) <- colnames(Y_real)
regional_rmse <- rmse(Y_real, Y_fitted)
plot(Y_real, Y_fitted) ; print(regional_rmse)  
```

### NR dataset

```{r}
northern_range_no_network_ZI <- ZIPLN(Abundance ~ 1 + PC1 + PC2 | 1,
                                      data = northern_range_data_pc)
```

```{r}
Y_real   <- northern_range_data_pc$Abundance
Y_fitted <- northern_range_no_network_ZI$fitted ; colnames(Y_fitted) <- colnames(Y_real)
northern_range_rmse <- rmse(Y_real, Y_fitted)
plot(Y_real, Y_fitted) ; print(northern_range_rmse)  
```

```{r}
fitting_results <- rbind(fitting_results, c("ZIPLN_PC1_PC2_ZIcol",
                                            round(regional_rmse, 3),
                                            round(northern_range_rmse, 3)))
```

## ZIPLN - Covariates effects in the ZI

### ZI PC1

#### regional dataset

```{r}
regional_no_network_ZI_PC1 <- ZIPLN(Abundance ~ 1 + PC1 + PC2 | 1 + PC1 ,
                                data = regional_data_pc)
```

```{r}
Y_real   <- regional_data_pc$Abundance
Y_fitted <- regional_no_network_ZI$fitted ; colnames(Y_fitted) <- colnames(Y_real)
regional_rmse <- rmse(Y_real, Y_fitted)
plot(Y_real, Y_fitted) ; print(regional_rmse)  
```

#### NR dataset

```{r}
northern_range_no_network_ZI <- ZIPLN(Abundance ~ 1 + PC1 + PC2  | 1 + PC1,
                                      data = northern_range_data_pc)
```

```{r}
Y_real   <- northern_range_data_pc$Abundance
Y_fitted <- northern_range_no_network_ZI$fitted ; colnames(Y_fitted) <- colnames(Y_real)
northern_range_rmse <- rmse(Y_real, Y_fitted)
plot(Y_real, Y_fitted) ; print(northern_range_rmse)  
```

```{r}
fitting_results <- rbind(fitting_results, c("ZIPLN_PC1_PC2_ZIcolPC1",
                                            round(regional_rmse, 3),
                                            round(northern_range_rmse, 3)))
```

### ZI PC2

#### regional dataset

```{r}
regional_no_network_ZI_PC1_PC2 <- ZIPLN(Abundance ~ 1 + PC1 + PC2 | 1 + PC1 + PC2,
                                        data = regional_data_pc)
```

```{r}
Y_real   <- regional_data_pc$Abundance
Y_fitted <- regional_no_network_ZI$fitted ; colnames(Y_fitted) <- colnames(Y_real)
regional_rmse <- rmse(Y_real, Y_fitted)
plot(Y_real, Y_fitted) ; print(regional_rmse)  
```

#### NR dataset

```{r}
# northern_range_no_network_ZI <- ZIPLN(Abundance ~ 1 + PC1 + PC2  | 1 + PC1 + PC2,
#                                       data = northern_range_data_pc)
```

```{r}
# Y_real   <- northern_range_data_pc$Abundance
# Y_fitted <- northern_range_no_network_ZI$fitted ; colnames(Y_fitted) <- colnames(Y_real)
# northern_range_rmse <- rmse(Y_real, Y_fitted)
# plot(Y_real, Y_fitted) ; print(northern_range_rmse)  
northern_range_rmse <- NA
```

```{r}
fitting_results <- rbind(fitting_results, c("ZIPLN_PC1_PC2_ZIcolPC1PC2",
                                            round(regional_rmse, 3),
                                            round(northern_range_rmse, 3)))
```

```{r}
print(fitting_results)
```

# Running PLN-network (no zero-inflation)

### Fitting the model

#### Regional dataset

```{r regional_model_no_ZI}
regional_no_ZI <- PLNnetwork(Abundance ~ 1 + PC1 + PC2,
                             data = regional_data_pc,
                             control = PLNnetwork_param(min_ratio = 0.05,
                                                        penalize_diagonal = FALSE))
```

```{r}
# regional_no_ZI_StARS <- regional_no_ZI$getBestModel("StARS")
regional_no_ZI_StARS <- regional_no_ZI$getModel(1.191822)
```

#### Northern Range dataset

```{r northern_range_model_no_ZI}
northern_range_no_ZI <- PLNnetwork(Abundance ~ 1 + PC1 + PC2,
                                  data = northern_range_data_pc,
                                  control = PLNnetwork_param(min_ratio = 0.05,
                                                             penalize_diagonal = FALSE))
```

```{r}
# northern_range_no_ZI_StARS <- northern_range_no_ZI$getBestModel("StARS")
northern_range_no_ZI_StARS <- northern_range_no_ZI$getModel(0.6133228)
```

### Goodness of fit

#### Regional dataset

```{r regional_fit}
Y_real   <- regional_data_pc$Abundance
Y_fitted <- regional_no_ZI_StARS$fitted ; colnames(Y_fitted) <- colnames(Y_real)
plot(Y_real, Y_fitted)
```

```{r}
print(rmse(Y_real, Y_fitted))
```

Focusing on the species with many 0s.

```{r regional_fit_zeros}
zero_threshold <- median(colSums(Y_real == 0))
zero_species   <- names(which(colSums(Y_real == 0) > zero_threshold & colSums(Y_real) < 400))
Y_zero         <- Y_real[, zero_species] ; Y_fitted_zero <- Y_fitted[, zero_species] 
plot(Y_zero, Y_fitted_zero)
```

```{r}
print(rmse(Y_zero, Y_fitted_zero))
```

#### Northern Range dataset

```{r nr_fit}
Y_real   <- northern_range_data$Abundance
Y_fitted <- northern_range_no_ZI_StARS$fitted ; colnames(Y_fitted) <- colnames(Y_real)
plot(Y_real, Y_fitted)
```

```{r}
print(rmse(Y_real, Y_fitted))
```

Focusing on the species with many 0s.

```{r nr_fit_zeros}
zero_threshold <- median(colSums(Y_real == 0))
zero_species   <- names(which(colSums(Y_real == 0) > zero_threshold & colSums(Y_real) < 400))
Y_zero         <- Y_real[, zero_species] ; Y_fitted_zero <- Y_fitted[, zero_species] 
plot(Y_zero, Y_fitted_zero)
```

```{r}
print(rmse(Y_zero, Y_fitted_zero))
```

### Resulting networks

#### Regional dataset

```{r}
plot(regional_no_ZI_StARS)
```

With a 0 penalty we would have:

```{r}
plot(regional_no_ZI$getModel(0))
```

#### Northern Range dataset

```{r}
plot(northern_range_no_ZI_StARS)
```

With a 0 penalty we would have:

```{r}
plot(northern_range_no_ZI$getModel(0))
```

### Robustness of the networks to the addition of covariates

#### Regional network

```{r}
plot(regional_no_ZI_StARS)
```

```{r}
regional_no_ZI_PC3 <- PLNnetwork(Abundance ~ 1 + PC1 + PC2 + PC3,
                                 data = regional_data_pc,
                                 control = PLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE))
# regional_no_ZI_PC3_StARS <- regional_no_ZI_PC3$getBestModel("StARS")
regional_no_ZI_PC3_StARS <- regional_no_ZI_PC3$getModel(1.15029)
plot(regional_no_ZI_PC3_StARS)
```

```{r}
regional_no_ZI_LONGLAT <- PLNnetwork(Abundance ~ 1 + LONGLAT,
                                 data = regional_data,
                                 control = PLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE))
# regional_no_ZI_LONGLAT_StARS <- regional_no_ZI_LONGLAT$getBestModel("StARS")
regional_no_ZI_LONGLAT_StARS <- regional_no_ZI_LONGLAT$getModel(1.153061)
plot(regional_no_ZI_LONGLAT_StARS)
```

```{r}
robustness(c(regional_no_ZI_StARS,
             regional_no_ZI_PC3_StARS,
             regional_no_ZI_LONGLAT_StARS),
           c("regional_no_ZI_StARS",
             "regional_no_ZI_PC3_StARS",
             "regional_no_ZI_LONGLAT_StARS" ))
```

#### Northern Range network

```{r}
plot(northern_range_no_ZI_StARS)
```

```{r}
northern_range_no_ZI_PC3 <- PLNnetwork(Abundance ~ 1 + PC1 + PC2 + PC3,
                                       data = northern_range_data_pc,
                                       control = PLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE))
# northern_range_no_ZI_PC3_StARS <- northern_range_no_ZI_PC3$getBestModel("StARS")
northern_range_no_ZI_PC3_StARS <- northern_range_no_ZI_PC3$getModel(0.5743538)
plot(northern_range_no_ZI_PC3_StARS)
```

```{r}
northern_range_no_ZI_LONGLAT <- PLNnetwork(Abundance ~ 1 + LONGLAT,
                                 data = northern_range_data,
                                 control = PLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE))
# northern_range_no_ZI_LONGLAT_StARS <- northern_range_no_ZI_LONGLAT$getBestModel("StARS")
northern_range_no_ZI_LONGLAT_StARS <- northern_range_no_ZI_LONGLAT$getModel(0.6260702)
plot(northern_range_no_ZI_LONGLAT_StARS)
```

```{r}
robustness(c(northern_range_no_ZI_StARS,
             northern_range_no_ZI_PC3_StARS,
             northern_range_no_ZI_LONGLAT_StARS),
           c("northern_range_no_ZI_StARS",
             "northern_range_no_ZI_PC3_StARS",
             "northern_range_no_ZI_LONGLAT_StARS" ))
```

# Running ZIPLN-network

### Fitting the model

#### Regional dataset

```{r regional_model_ZI}
regional_model<- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2,
                              data = regional_data_pc,
                              control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                           penalize_diagonal = FALSE),
                               zi = "col")
```

```{r}
# regional_model_StARS <- regional_model$getBestModel("StARS")
regional_model_StARS <- regional_model$getModel(0.07299546)
```

#### Northern Range dataset

```{r northern_range_model_ZI}
northern_range_model <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2,
                                     data = northern_range_data_pc,
                                     control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                                 penalize_diagonal = FALSE),
                                     zi = "col")
```

```{r}
# northern_range_model_StARS <- northern_range_model$getBestModel("StARS")
northern_range_model_StARS <- northern_range_model$getModel(0.08051953)
```

### Goodness of fit

#### Regional dataset

```{r regional_fit_ZI}
Y_real   <- regional_data_pc$Abundance
Y_fitted <- regional_model_StARS$fitted ; colnames(Y_fitted) <- colnames(Y_real)
plot(Y_real, Y_fitted)
```

```{r}
print(rmse(Y_real, Y_fitted))
```

Focusing on the species with many 0s.

```{r regional_fit_zeros_ZI}
zero_threshold <- median(colSums(Y_real == 0))
zero_species   <- names(which(colSums(Y_real == 0) > zero_threshold & colSums(Y_real) < 400))
Y_zero         <- Y_real[, zero_species] ; Y_fitted_zero <- Y_fitted[, zero_species] 
plot(Y_zero, Y_fitted_zero)
```

```{r}
print(rmse(Y_zero, Y_fitted_zero))
```

#### Northern Range dataset

```{r nr_fit_ZI}
Y_real   <- northern_range_data$Abundance
Y_fitted <- northern_range_model_StARS$fitted ; colnames(Y_fitted) <- colnames(Y_real)
plot(Y_real, Y_fitted)
```

```{r}
print(rmse(Y_real, Y_fitted))
```

Focusing on the species with many 0s.

```{r nr_fit_zeros_ZI}
zero_threshold <- median(colSums(Y_real == 0))
zero_species   <- names(which(colSums(Y_real == 0) > zero_threshold & colSums(Y_real) < 400))
Y_zero         <- Y_real[, zero_species] ; Y_fitted_zero <- Y_fitted[, zero_species] 
plot(Y_zero, Y_fitted_zero)
```

```{r}
print(rmse(Y_zero, Y_fitted_zero))
```

### Resulting networks

#### Regional dataset

```{r}
plot(regional_model_StARS)
```

With a 0 penalty we would have:

```{r}
plot(regional_model$getModel(0))
```

Remark: we see that several associations are the same as those found with LONGLAT (Rhamquel - Coryriis, Poecreti - Anabhart, Sicypunc - Eleopiso).

#### Northern Range dataset

```{r}
plot(northern_range_model_StARS)
```

Remark: again, several associations are the same as those found with LONGLAT (Hoplmala - Anabhart, Rhamquel - Anabhart, Poecreti - Andipulc...).

With a 0 penalty we would have:

```{r}
plot(northern_range_model$getModel(0))
```

### Robustness of the networks to the addition of covariates

#### Regional network

```{r}
plot(regional_model_StARS)
```

```{r}
regional_PC3 <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 + PC3,
                             data = regional_data_pc,
                             control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE),
                             zi = "col")
# regional_PC3_StARS <- regional_PC3$getBestModel("StARS")
regional_PC3_StARS <- regional_PC3$getModel(0.08177813)
# plot(regional_PC3_StARS)
```

```{r}
regional_LONGLAT <- ZIPLNnetwork(Abundance ~ 1 + LONGLAT,
                                 data = regional_data,
                                 control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE),
                                 zi = "col")
# regional_LONGLAT_StARS <- regional_LONGLAT$getBestModel("StARS")
regional_LONGLAT_StARS <- regional_LONGLAT$getModel(0.06633124)
plot(regional_LONGLAT_StARS)
```

```{r}
regional_ziPC1 <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 | 1 + PC1,
                             data = regional_data_pc,
                             control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE))
# regional_ziPC1_StARS <- regional_ziPC1$getBestModel("StARS")
regional_ziPC1_StARS <- regional_ziPC1$getModel(0.03697966)
plot(regional_ziPC1_StARS)
```

```{r}
regional_ziPC1PC2 <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 | 1 + PC1 + PC2,
                             data = regional_data_pc,
                             control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE))
# regional_ziPC1PC2_StARS <- regional_ziPC1PC2$getBestModel("StARS")
regional_ziPC1PC2_StARS <- regional_ziPC1PC2$getModel(0.03828855)
# plot(regional_ziPC1_StARS)
```

```{r}
robustness(c(regional_model_StARS,
             regional_PC3_StARS,
             regional_LONGLAT_StARS,
             regional_ziPC1_StARS,
             regional_ziPC1PC2_StARS),
           c("regional_StARS",
             "regional_PC3_StARS",
             "regional_LONGLAT_StARS",
             "regional_ziPC1_StARS",
             "regional_ziPC1PC2_StARS" ))
```

#### Northern Range network

```{r}
plot(northern_range_model_StARS)
```

```{r}
# northern_range_PC3 <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 + PC3,
#                              data = northern_range_data_pc,
#                              control = ZIPLNnetwork_param(min_ratio = 0.05,
#                                                             penalize_diagonal = FALSE),
#                              zi = "col")
# regional_PC3_StARS <- regional_PC3$getBestModel("StARS")
# regional_PC3_StARS <- regional_PC3$getModel(0.08177813)
# plot(regional_PC3_StARS)
```

```{r}
northern_range_LONGLAT <- ZIPLNnetwork(Abundance ~ 1 + LONGLAT,
                                               data = northern_range_data,
                                               control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE),
                                               zi = "col")
# northern_range_LONGLAT_StARS <- northern_range_LONGLAT$getBestModel("StARS")
northern_range_LONGLAT_StARS <- northern_range_LONGLAT$getModel(0.08978182)
# plot(regional_LONGLAT_StARS)
```

```{r}
northern_range_ziPC1 <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 | 1 + PC1,
                                     data = northern_range_data_pc,
                                     control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                            penalize_diagonal = FALSE))
# northern_range_ziPC1_StARS <- northern_range_ziPC1$getBestModel("StARS")
northern_range_ziPC1_StARS <- northern_range_ziPC1$getModel(0.04591576)
plot(northern_range_ziPC1_StARS)
```

```{r}
# Error in if (parameters_converged | objective_converged)
# northern_range_ziPC1PC2 <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 | 1 + PC1 + PC2,
#                              data = northern_range_data_pc,
#                              control = ZIPLNnetwork_param(min_ratio = 0.05,
#                                                             penalize_diagonal = FALSE))
# northern_range_ziPC1PC2_StARS <- northern_range_ziPC1PC2$getBestModel("StARS")
# northern_range_ziPC1PC2_StARS   <- northern_range_ziPC1PC2$getModel(0.03828855)
# plot(northern_range_ziPC1PC2_StARS)
```

```{r}
robustness(c(northern_range_model_StARS,
             northern_range_LONGLAT_StARS,
             northern_range_ziPC1_StARS),
           c("northern_range_StARS",
             "northern_range_LONGLAT_StARS",
             "northern_range_ziPC1_StARS"))
```

# Accounting for sampling date in the Northern Range dataset

## Formatting date as appropriate covariates

```{r}
date_covar <- data.frame("Year"  = as.factor(NORTHERN_RANGE_DATASET_COVARIATES$YEAR),
                         "Month" = as.factor(NORTHERN_RANGE_DATASET_COVARIATES$MONTH))
northern_range_date_pc_covar  <- cbind(date_covar, pca_covar_northern_range)
rownames(northern_range_date_pc_covar) <- rownames(NORTHERN_RANGE_DATASET_COVARIATES)
northern_range_data_date_pc <- prepare_data(counts     = NORTHERN_RANGE_DATASET_COUNTS,
                                            covariates = northern_range_date_pc_covar)
```

## Re-running PLN-network with the date covariates

```{r}
northern_range_no_ZI_model_year <- PLNnetwork(Abundance ~ 1 + PC1 + PC2 + Year,
                                              data = northern_range_data_date_pc,
                                              control = PLNnetwork_param(min_ratio = 0.05,
                                                           penalize_diagonal = FALSE))
# northern_range_no_ZI_model_year_StARS <- northern_range_no_ZI_model_year$getBestModel("StARS")
northern_range_no_ZI_model_year_StARS <- northern_range_no_ZI_model_year$getModel(0.5421913)
# plot(northern_range_no_ZI_model_year_StARS)
```

```{r}
northern_range_no_ZI_model_year_month <- PLNnetwork(Abundance ~ 1 + PC1 + PC2 + Year +  Month,
                                                    data = northern_range_data_date_pc,
                                                    control = PLNnetwork_param(min_ratio = 0.05,
                                                              penalize_diagonal = FALSE))
# northern_range_no_ZI_model_year_month_StARS <- northern_range_no_ZI_model_year_month$getBestModel("StARS")
northern_range_no_ZI_model_year_month_StARS <- northern_range_no_ZI_model_year_month$getModel(0.3764233)
# plot(northern_range_model_year_month_StARS)
```

## Re-running ZIPLN-network with the date covariates

```{r}
northern_range_model_year <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 + Year,
                                          data = northern_range_data_date_pc,
                                          control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                           penalize_diagonal = FALSE),
                                          zi = "col")
# northern_range_model_year_StARS <- northern_range_model_year$getBestModel("StARS")
northern_range_model_year_StARS <- northern_range_model_year$getModel(0.07216102)
plot(northern_range_model_year_StARS)
```

Remark: at first sight, it appears that adding the Year as a covariate does not significantly change the network structure.
```{r}
northern_range_model_year_ZIyear <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 + Year | 1 + Year,
                                          data = northern_range_data_date_pc,
                                          control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                           penalize_diagonal = FALSE))
# northern_range_model_year_ZIyear_StARS <- northern_range_model_year_ZIyear$getBestModel("StARS")
northern_range_model_year_ZIyear_StARS <- northern_range_model_year_ZIyear$getModel(0.07185885)
plot(northern_range_model_year_ZIyear_StARS)
```

```{r}
# northern_range_model_month <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 + Month,
#                                           data = northern_range_data_date_pc,
#                                           control = ZIPLNnetwork_param(min_ratio = 0.05,
#                                                            penalize_diagonal = FALSE),
#                                           zi = "col")
# Error in if (parameters_converged | objective_converged) { : 
# missing value where TRUE/FALSE needed
```

```{r}
northern_range_model_year_month <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 + Year +  Month,
                                               data = northern_range_data_date_pc,
                                               control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                           penalize_diagonal = FALSE),
                                               zi = "col")
# northern_range_model_year_month_StARS <- northern_range_model_year_month$getBestModel("StARS")
northern_range_model_year_month_StARS <- northern_range_model_year_month$getModel(0.05944257)
plot(northern_range_model_year_month_StARS)
```

Remark: one significant change is the disappearing of the Rhamquel - Anabhart association. A few associations lose strength (Hoplamala - Rhamquel) while others seem stronger (Ancimara - Hyporobi).

```{r}
northern_range_model_ZImonth <- ZIPLNnetwork(Abundance ~ 1 + PC1 + PC2 | 1 +  Month,
                                          data = northern_range_data_date_pc,
                                          control = ZIPLNnetwork_param(min_ratio = 0.05,
                                                           penalize_diagonal = FALSE))
# northern_range_model_ZImonth_StARS <- northern_range_model_ZImonth$getBestModel("StARS")
northern_range_model_ZImonth_StARS <- northern_range_model_ZImonth$getModel(0.08764195)
plot(northern_range_model_ZImonth_StARS)
```

# Comparison of the networks with / without ZI

```{r}
robustness(c(regional_no_ZI_StARS,
             regional_no_ZI_PC3_StARS,
             regional_no_ZI_LONGLAT_StARS,
             regional_model_StARS,
             regional_PC3_StARS,
             regional_LONGLAT_StARS,
             regional_ziPC1_StARS,
             regional_ziPC1PC2_StARS),
           c("regional_no_ZI_StARS",
             "regional_no_ZI_PC3_StARS",
             "regional_no_ZI_LONGLAT_StARS",
             "regional_StARS",
             "regional_PC3_StARS",
             "regional_LONGLAT_StARS",
             "regional_ziPC1_StARS",
             "regional_ziPC1PC2_StARS" ))
```

```{r}
robustness(c(northern_range_no_ZI_StARS,
             northern_range_no_ZI_LONGLAT_StARS,
             northern_range_no_ZI_model_year_StARS,
             northern_range_no_ZI_model_year_month_StARS,
             northern_range_model_StARS,
             northern_range_LONGLAT_StARS,
             northern_range_ziPC1_StARS,
             northern_range_model_year_StARS,
             northern_range_model_year_month_StARS),
           c("northern_range_no_ZI_StARS",
             "northern_range_no_ZI_LONGLAT_StARS",
             "northern_range_no_ZI_model_year_StARS",
             "northern_range_no_ZI_model_year_month_StARS",
             "northern_range_StARS",
             "northern_range_LONGLAT_StARS",
             "northern_range_ziPC1_StARS",
             "northern_range_model_year_StARS",
             "northern_range_model_year_month_StARS"))
```

# Comparison with Hmsc package
```{r}
Y_regional <- regional_data$Abundance
Y_NR       <- northern_range_data$Abundance
```

## With LONGLAT
### Regional data
```{r}
X_regional_LONGLAT    <- REGIONAL_DATASET_COVARIATES %>%  select(c("LONGLAT"))

nChains = 2
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
n = nrow(Y_regional)

studyDesign = data.frame(sample = as.factor(1:n))
rL = HmscRandomLevel(units = studyDesign$sample)
m_regional_LONGLAT = Hmsc(Y = Y_regional, XData = X_regional_LONGLAT, XFormula = ~LONGLAT,
studyDesign = studyDesign, ranLevels = list(sample = rL), distr = "lognormal poisson")
m_regional_LONGLAT  = sampleMcmc(m_regional_LONGLAT , thin = thin, samples = samples, transient = transient,
nChains = nChains, nParallel = nChains, verbose = verbose)
```

```{r}
preds = computePredictedValues(m_regional_LONGLAT )
evaluateModelFit(hM = m_regional_LONGLAT , predY = preds)
```

```{r}
OmegaCor = computeAssociations(m_regional_LONGLAT)
supportLevel = 0.95
Sigma_regional_LONGLAT = ((OmegaCor[[1]]$support>supportLevel)
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean
corrplot(Sigma_regional_LONGLAT, method = "color",
col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", m_regional_LONGLAT $rLNames[1]), mar=c(0,0,1,0))
```
```{r}
plot_network(Sigma_regional_LONGLAT * 0.1)
```

### Northern Range data
```{r}
X_NR_LONGLAT    <- northern_range_data %>%  select(c("LONGLAT"))
nChains = 2
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
n = nrow(Y_NR)

studyDesign = data.frame(sample = as.factor(1:n))
rL = HmscRandomLevel(units = studyDesign$sample)
m_NR_LONGLAT = Hmsc(Y = Y_NR, XData = X_NR_LONGLAT, XFormula = ~LONGLAT,
studyDesign = studyDesign, ranLevels = list(sample = rL), distr = "lognormal poisson")
m_NR_LONGLAT = sampleMcmc(m_NR_LONGLAT, thin = thin, samples = samples, transient = transient,
nChains = nChains, nParallel = nChains, verbose = verbose)
```

```{r}
preds = computePredictedValues(m_NR_LONGLAT)
evaluateModelFit(hM = m_NR_LONGLAT, predY = preds)
```

```{r}
OmegaCor = computeAssociations(m_NR_LONGLAT)
supportLevel = 0.95
Sigma_NR_LONGLAT = ((OmegaCor[[1]]$support>supportLevel)
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean
corrplot(Sigma_NR_LONGLAT, method = "color",
col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", m_NR_LONGLAT$rLNames[1]), mar=c(0,0,1,0))
```

```{r}
plot_network(Sigma_NR_LONGLAT*0.1)
```

## With PC1 + PC1
### Regional data
```{r}
X_regional_PC1_PC2    <- regional_data_pc %>%  select(c("PC1", "PC2"))

nChains = 2
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
n = nrow(Y_regional)

studyDesign = data.frame(sample = as.factor(1:n))
rL = HmscRandomLevel(units = studyDesign$sample)
m_regional_PC1_PC2 = Hmsc(Y = Y_regional, XData = X_regional_PC1_PC2, XFormula = ~ PC1 + PC2,
studyDesign = studyDesign, ranLevels = list(sample = rL), distr = "lognormal poisson")
m_regional_PC1_PC2  = sampleMcmc(m_regional_PC1_PC2, thin = thin, samples = samples, transient = transient,
nChains = nChains, nParallel = nChains, verbose = verbose)
```

```{r}
preds = computePredictedValues(m_regional_PC1_PC2 )
evaluateModelFit(hM = m_regional_PC1_PC2 , predY = preds)
```

```{r}
OmegaCor = computeAssociations(m_regional_PC1_PC2)
supportLevel = 0.95
Sigma_regional_PC1_PC2 = ((OmegaCor[[1]]$support>supportLevel)
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean
corrplot(Sigma_regional_PC1_PC2, method = "color",
col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", m_regional_LONGLAT $rLNames[1]), mar=c(0,0,1,0))
```

```{r}
plot_network(Sigma_regional_PC1_PC2 * 0.1)
```

### Northern Range data
```{r}
X_NR_PC1_PC2    <- northern_range_data_pc %>%  select(c("PC1", "PC2"))

nChains = 2
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
n = nrow(Y_NR)

studyDesign = data.frame(sample = as.factor(1:n))
rL = HmscRandomLevel(units = studyDesign$sample)
m_NR_PC1_PC2 = Hmsc(Y = Y_NR, XData = X_NR_PC1_PC2, XFormula = ~PC1 + PC2,
studyDesign = studyDesign, ranLevels = list(sample = rL), distr = "lognormal poisson")
m_NR_PC1_PC2 = sampleMcmc(m_NR_PC1_PC2, thin = thin, samples = samples, transient = transient,
nChains = nChains, nParallel = nChains, verbose = verbose)
```

```{r}
preds = computePredictedValues(m_NR_PC1_PC2)
evaluateModelFit(hM = m_NR_PC1_PC2, predY = preds)
```

```{r}
OmegaCor = computeAssociations(m_NR_PC1_PC2)
supportLevel = 0.95
Sigma_NR_PC1_PC2 = ((OmegaCor[[1]]$support>supportLevel)
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean
corrplot(Sigma_NR_PC1_PC2, method = "color",
col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", m_NR_PC1_PC2$rLNames[1]), mar=c(0,0,1,0))
```

```{r}
plot_network(Sigma_NR_PC1_PC2 * 0.1)
```

## With PC1 + PC2 + dates
### Just year
```{r}
X_NR_PC1_PC2_year    <- northern_range_data_date_pc %>%  select(c("PC1", "PC2", "Year"))

nChains = 2
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
n = nrow(Y_NR)

studyDesign = data.frame(sample = as.factor(1:n))
rL = HmscRandomLevel(units = studyDesign$sample)
m_NR_PC1_PC2_year = Hmsc(Y = Y_NR, XData = X_NR_PC1_PC2_year, XFormula = ~PC1 + PC2 + Year,
studyDesign = studyDesign, ranLevels = list(sample = rL), distr = "lognormal poisson")
m_NR_PC1_PC2_year = sampleMcmc(m_NR_PC1_PC2_year, thin = thin, samples = samples, transient = transient,
nChains = nChains, nParallel = nChains, verbose = verbose)
```

```{r}
preds = computePredictedValues(m_NR_PC1_PC2_year)
evaluateModelFit(hM = m_NR_PC1_PC2_year, predY = preds)
```

```{r}
OmegaCor = computeAssociations(m_NR_PC1_PC2_year)
supportLevel = 0.95
Sigma_NR_PC1_PC2_Year = ((OmegaCor[[1]]$support>supportLevel)
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean
corrplot(Sigma_NR_PC1_PC2_Year, method = "color",
col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", m_NR_PC1_PC2_year$rLNames[1]), mar=c(0,0,1,0))
```

```{r}
plot_network(Sigma_NR_PC1_PC2_Year*0.1)
```

### Just month
```{r}
X_NR_PC1_PC2_month     <- northern_range_data_date_pc %>%  select(c("PC1", "PC2", "Month"))

nChains = 2
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
n = nrow(Y_NR)

studyDesign = data.frame(sample = as.factor(1:n))
rL = HmscRandomLevel(units = studyDesign$sample)
m_NR_PC1_PC2_month = Hmsc(Y = Y_NR, XData = X_NR_PC1_PC2_month, XFormula = ~ PC1 + PC2 + Month,
studyDesign = studyDesign, ranLevels = list(sample = rL), distr = "lognormal poisson")
m_NR_PC1_PC2_month = sampleMcmc(m_NR_PC1_PC2_month, thin = thin, samples = samples, transient = transient,
nChains = nChains, nParallel = nChains, verbose = verbose)
```

```{r}
preds = computePredictedValues(m_NR_PC1_PC2_month)
evaluateModelFit(hM = m_NR_PC1_PC2_month, predY = preds)
```

```{r}
OmegaCor = computeAssociations(m_NR_PC1_PC2_month)
supportLevel = 0.95
Sigma_NR_PC1_PC2_Month = ((OmegaCor[[1]]$support>supportLevel)
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean
corrplot(Sigma_NR_PC1_PC2_Month, method = "color",
col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", m_NR_PC1_PC2_month$rLNames[1]), mar=c(0,0,1,0))
```

```{r}
# Omega <- solve(Sigma_NR_PC1_PC2_Month)
# partial_corr <- - Omega / tcrossprod(sqrt(diag(Omega))); diag(partial_corr) <- 1
# plot_network(partial_corr)
plot_network(Sigma_NR_PC1_PC2_Month * 0.1)
```

### Year + Month
```{r}
X_NR_PC1_PC2_year_month     <- northern_range_data_date_pc %>%  select(c("PC1", "PC2", "Year", "Month"))
nChains = 2
thin = 10
samples = 1000
transient = 500*thin
verbose = 0
n = nrow(Y_NR)

studyDesign = data.frame(sample = as.factor(1:n))
rL = HmscRandomLevel(units = studyDesign$sample)
m_NR_PC1_PC2_year_month = Hmsc(Y = Y_NR, XData = X_NR_PC1_PC2_year_month, XFormula = ~ PC1 + PC2 + Year + Month,
studyDesign = studyDesign, ranLevels = list(sample = rL), distr = "lognormal poisson")
m_NR_PC1_PC2_year_month = sampleMcmc(m_NR_PC1_PC2_year_month, thin = thin, samples = samples, transient = transient,
nChains = nChains, nParallel = nChains, verbose = verbose)
```

```{r}
preds = computePredictedValues(m_NR_PC1_PC2_year_month)
evaluateModelFit(hM = m_NR_PC1_PC2_year_month, predY = preds)
```

```{r}
OmegaCor = computeAssociations(m_NR_PC1_PC2_year_month)
supportLevel = 0.95
Sigma_NR_PC1_PC2_Year_Month = ((OmegaCor[[1]]$support>supportLevel)
+ (OmegaCor[[1]]$support<(1-supportLevel))>0)*OmegaCor[[1]]$mean
corrplot(Sigma_NR_PC1_PC2_Year_Month, method = "color",
col = colorRampPalette(c("blue","white","red"))(200),
title = paste("random effect level:", m_NR_PC1_PC2_year_month$rLNames[1]), mar=c(0,0,1,0))
```

```{r}
plot_network(Sigma_NR_PC1_PC2_Year_Month*0.1)
```

## Comparing the HMSC networks together
### Regional network
```{r}
res <- robustness(list(Sigma_regional_LONGLAT,
                    Sigma_regional_PC1_PC2),
                  c("Sigma_regional_LONGLAT",
                    "Sigma_regional_PC1_PC2"),
                  matrices_input = TRUE
                  )
print(res)
```

### Northern Range network
```{r}
res <- robustness(list(Sigma_NR_LONGLAT,
                    Sigma_NR_PC1_PC2,
                    Sigma_NR_PC1_PC2_Year,
                    Sigma_NR_PC1_PC2_Month,
                    Sigma_NR_PC1_PC2_Year_Month),
                  c("Sigma_NR_LONGLAT",
                    "Sigma_NR_PC1_PC2",
                    "Sigma_NR_PC1_PC2_Year",
                    "Sigma_NR_PC1_PC2_Month",
                    "Sigma_NR_PC1_PC2_Year_Month"),
                  matrices_input = TRUE
                  )
print(res)
```

## Comparing the HMSC networks with the PLN ones
### Regional network
```{r}
res <- robustness(list(Sigma_regional_LONGLAT,
                       Sigma_regional_PC1_PC2,
                       as.matrix(regional_no_ZI_StARS$latent_network("partial_cor")),
                       as.matrix(regional_no_ZI_PC3_StARS$latent_network("partial_cor")),
                       as.matrix(regional_no_ZI_LONGLAT_StARS$latent_network("partial_cor")),
                       as.matrix(regional_model_StARS$latent_network("partial_cor")),
                       as.matrix(regional_PC3_StARS$latent_network("partial_cor")),
                       as.matrix(regional_LONGLAT_StARS$latent_network("partial_cor")),
                       as.matrix(regional_ziPC1_StARS$latent_network("partial_cor")),
                       as.matrix(regional_ziPC1PC2_StARS$latent_network("partial_cor"))),
                  c("HMSC_Sigma_regional_LONGLAT",
                    "HMSC_Sigma_regional_PC1_PC2",
                    "regional_no_ZI_StARS",
                    "regional_no_ZI_PC3_StARS",
                    "regional_no_ZI_LONGLAT_StARS",
                    "regional_model_StARS",
                    "regional_PC3_StARS",
                    "regional_LONGLAT_StARS",
                    "regional_ziPC1_StARS",
                    "regional_ziPC1PC2_StARS"),
                  matrices_input = TRUE
                  )
print(res)
```

### Northern Range network
```{r}
res <- robustness(list(Sigma_NR_LONGLAT,
                       Sigma_NR_PC1_PC2,
                       Sigma_NR_PC1_PC2_Year,
                       Sigma_NR_PC1_PC2_Month,
                       Sigma_NR_PC1_PC2_Year_Month,
                       as.matrix(northern_range_no_ZI_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_no_ZI_LONGLAT_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_no_ZI_model_year_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_no_ZI_model_year_month_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_model_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_LONGLAT_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_ziPC1_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_model_year_StARS$latent_network("partial_cor")),
                       as.matrix(northern_range_model_year_month_StARS$latent_network("partial_cor"))),
                  c("HSMC_Sigma_NR_LONGLAT",
                    "HSMC_Sigma_NR_PC1_PC2",
                    "HSMC_Sigma_NR_PC1_PC2_Year",
                    "HSMC_Sigma_NR_PC1_PC2_Month",
                    "HSMC_Sigma_NR_PC1_PC2_Year_Month",
                    "northern_range_no_ZI_StARS",
                    "northern_range_no_ZI_LONGLAT_StARS",
                    "northern_range_no_ZI_model_year_StARS",
                    "northern_range_no_ZI_model_year_month_StARS",
                    "northern_range_model_StARS",
                    "northern_range_LONGLAT_StARS",
                    "northern_range_ziPC1_StARS",
                    "northern_range_model_year_StARS",
                    "northern_range_model_year_month_StARS"),
                  matrices_input = TRUE
                  )

print(res)
```